Modeling Macroeconomic Phenomena \\with Multi-Agent Reinforcement Learning

### 2019-12-06 paper/paper.tex

Recall that the loss function for the REINFORCE algorithm is of the following form:
\[
  \mathcal{L} = -\sum_{t = 1}^T \left[\log \pi_\theta(s_t, a_t) \left( \sum_{i = t}^n \gamma^{i - t}r_i \right) \right].  
\]

In particular, as the policy converges, the probability of taking a given action $a_t$ from state $s_t$ should approach one, and so the $\log \pi_\theta$ term will approach zero. Therefore, convergence corresponds to the losses approaching zero. This occurs in our preliminary simulation, as can be seen in Figure \ref{plt:losstozero}.

\begin{figure}[h]
\begin{tikzpicture}[every mark/.append style={mark size=1pt}]
  \begin{axis}[
    xlabel=Episode,
    ylabel=Loss,
    legend style={at={(1,0)}, anchor=south east}
  ]
  \addplot table [y=fl, x=i]{prelims.dat};
  \addlegendentry{Firm}
  \addplot table [y=pl, x=i]{prelims.dat};
  \addlegendentry{Human (Mean)}
  \end{axis}
\end{tikzpicture}
\caption{The agents' losses converge to zero over time. This plot illustrates the loss of firm agents vs. human agents in a simulation with one firm agent and five human agents. The loss for the human agents is the average of all five human agents' losses. This simulation uses the REINFORCE learning algorithm, and indicates that the algorithm is succcessfully converging to a policy within 100 episodes.}
\label{plt:losstozero}
\end{figure}

>>>>

Recall that the gradient the REINFORCE algorithm is of the following form:
\[
  \nabla_{\theta} J(\theta) = -\sum_{t = 1}^T \left[\nabla_{\theta} \log \pi_\theta(s_t, a_t) \left( \sum_{i = t}^n \gamma^{i - t}r_i \right) \right].  
\]

In our training, we designate a proxy for the loss to reflect this gradient:
\[
  J'(\theta) = -\sum_{t = 1}^T \left[\log \pi_\theta(s_t, a_t) \left( \sum_{i = t}^n \gamma^{i - t}r_i \right) \right].  
\]
Note that, as the policy converges, the probability of taking a given action $a_t$ from state $s_t$ should approach one, and so the $\log \pi_\theta$ term will approach zero. Therefore, convergence corresponds to the (proxy) losses approaching zero. This occurs in our preliminary simulation, as can be seen in Figure \ref{plt:losstozero}.

\begin{figure}[h]
\begin{tikzpicture}[every mark/.append style={mark size=1pt}]
  \begin{axis}[
    xlabel=Episode,
    ylabel=Loss (Proxy),
    legend style={at={(1,0)}, anchor=south east}
  ]
  \addplot table [y=fl, x=i]{prelims.dat};
  \addlegendentry{Firm}
  \addplot table [y=pl, x=i]{prelims.dat};
  \addlegendentry{Human (Mean)}
  \end{axis}
\end{tikzpicture}
\caption{The agents' losses converge to zero over time. This plot illustrates the loss of firm agents vs. human agents in a simulation with one firm agent and five human agents. The loss for the human agents is the average of all five human agents' losses. This simulation uses the REINFORCE learning algorithm, and indicates that the algorithm is succcessfully converging to a policy within 100 episodes.}
\label{plt:losstozero}
\end{figure}
